---
title: "Final Lab_10"
author: "Bhumika Subedi-Giri"
date: "4/1/2020"
output: html_document
---
### The data used for graphs below are taken from Jonh Hopkins Github site
```{r, warnings=FALSE, message=FALSE, echo=FALSE }
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
```

### This graph shows number of Death cases of COVID-19 around the world on March 23rd, 2020.
```{r, warnings=FALSE, message=FALSE, echo=FALSE}  
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-23-2020.csv")) %>% 
  rename(Long = "Long_") 
    
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='green', fill='green', alpha = 0.5) +
    labs(title = 'World COVID-19 Deaths cases',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

### The graph below shows the number of comfirmed cases of COVID-19 on 48 states of the US on 03/31/2020. However, it does not incude Hawaii and Alaska and other US Territories. Also, some US state entries have a Lat and Long of Zero, these are filtered, and not shown. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)

ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "grey", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='maroon', fill='maroon', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)

```

### The graph below shows the same number of Death cases as above graph but it is much better to look at. This was created based on the example by Anisa Dhana.
```{r, warning=FALSE, message=FALSE, echo=FALSE}
mybreaks <- c(1, 100, 1000, 10000, 10000)

ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "Black", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.8) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

### The graph below shows the number of Death cases of COVID-19 in the US on March 28th, 2020. This graph uses one color gradient. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-28-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  mutate(Province_State = tolower(Province_State))

# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))

# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Deaths), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradient(trans = "log10") +
  labs(title = "COVID-19 Deaths in the US'")
```

### This graphs shows the number of Death cases of COVID-19 in the 50 States of the US. Here, the use of different colors allows us to differntiate the higher and lower number of death cases in different States faster as compared to the graph above. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Deaths), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradientn(colours = 
                         wes_palette("GrandBudapest2", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Deaths in the US'")
```

### The graph below shows the number of Death cases of COVID-19 on March 28th, 2020 in Connecticut. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-28-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Connecticut") %>% 
  group_by(Admin2) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  mutate(Admin2 = tolower(Admin2))

us <- map_data("state")
ct_us <- subset(us, region == "connecticut ")
counties <- map_data("county")
ct_county <- subset(counties, region == "connecticut")

state_join <- left_join(ct_county, daily_report, by = c("subregion" = "Admin2")) 

# plot state map
ggplot(data = ct_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Deaths), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("Cavalcanti1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Deaths in Connecticut '")
```

### Interactive maps
### This graphs is an interactive graph. It can tell you the exact number of death cases in each county of Connecticut, as you move your pointer on the Map.
```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(plotly)

ggplotly(
  ggplot(data = ct_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Deaths), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Cavalcanti1", 100, type = "continuous")) +
  ggtitle("COVID-19 Deaths Cases in CT") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

### Preparing time series data
```{r, warning=FALSE, message=FALSE, echo=FALSE}
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

# Let's get the times series data for deaths

time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")

time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

# Create Keys 

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))

time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)

# Reformat the data
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```

```{r warnings=FALSE, message=FALSE, echo=FALSE}
library(ggplot2)
library(gganimate)
library(transformr)
theme_set(theme_bw())
```

### Animation of Death cases of COVID-19 in China, South Korea, Italy and US. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Italy","US")) 

p <- ggplot(data_time, aes(x = Date,  y = Deaths, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Deaths COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
    
animate(p, renderer = gifski_renderer(), end_pause = 15)
```

### Adaption of animation code from Prof. Chris Sutherland. This animation gives us a visual of Death cases of COVID-19 around the world. It also shows which countries are affected by COVID-19 the most. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
           rename(Province_State= "Province/State", Country_Region = "Country/Region") %>%
           pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                  names_to = "Date", values_to = "Deaths") %>%
           mutate(Date = mdy(Date) - days(1),
                  Place = paste(Lat,Long,sep="_")) %>%
# Summarizes state and province information
             group_by(Place,Date) %>%
           summarise(cumulative_cases = ifelse(sum(Deaths)>0,
                     sum(Deaths),NA_real_),
                     Lat = mean(Lat),
                     Long = mean(Long)) %>%
           mutate(Pandemic_day = as.numeric(Date - min(Date)))

world <- ggplot(covid,aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
                 borders("world", colour = "gray50", fill = "grey90") +
                 theme_bw() +
                 geom_point(color='maroon', alpha = .5) +
                 labs(title = 'Pandemic Day: {frame}',x = '', y = '',
                      size="Cases (x1000))") +
                 theme(legend.position = "right") +
                 coord_fixed(ratio=1.3)+
                 transition_time(Date) +
                 enter_fade()

animate(world, renderer = gifski_renderer(), end_pause = 30)
```

